---
title: "Plot severity data from China"
author: "Katy Gaythorpe and Natsuko Imai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_width: 7
    fig_height: 5
    fig_caption: true
    highlight: "tango"
    reference_docx: word_style.docx
    df_print: "kable"
---

```{r set up }

 knitr::opts_chunk$set(echo = FALSE, fig.path = "figures/", fig.ext = "png", 
                       fig.height = 8)

## library(orderly)
## orderly::orderly_test_start("B2_plot_severity_china")
## orderly::orderly_run("B2_plot_severity_china")

```

Import data.
```{r }

df_symptom <- read.csv("severity_data_China_papers.csv",
                       stringsAsFactors = FALSE)

```

# General severity and comorbidity plots

```{r proportional_symptoms_all,  fig.cap="Proportion of all symptoms of each type."}
df_symptom %>%
  filter(grouped_symptoms != "any_symptom") %>%
  group_by(grouped_symptoms) %>%
  summarise(numerator = sum(symptom_numerator, na.rm = TRUE), 
            denominator = sum(denominator, na.rm = TRUE) ) %>%
  mutate(proportion_symptom= numerator/sum(denominator, na.rm = T),
         proportion_symptom_low = Hmisc::binconf(numerator,sum(denominator,
                                                               na.rm=T) )[,2],
         proportion_symptom_high = Hmisc::binconf(numerator,sum(denominator,
                                                                na.rm = T) )[,3])%>%

ggplot() + 
  geom_col(aes(x = reorder(grouped_symptoms, -proportion_symptom), 
                                  y = proportion_symptom,
                                  fill = proportion_symptom)) +
  geom_errorbar(aes(x = reorder(grouped_symptoms, -proportion_symptom), 
                    ymin = proportion_symptom_low,
                    ymax = proportion_symptom_high))+
  theme_pubr() +
  scale_fill_viridis_c()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        legend.position = "none") +
  labs(x = "Symptoms", y = "Proportion of all people reporting each symptom")

```

```{r proportional_symptoms_by_severity, fig.cap="Proportion of all symptoms of each type structured by severity."}
# plotting symptom occurrence by severity
tmp <- df_symptom %>%
  filter(grouped_symptoms != "any_symptom") %>%
  group_by(grouped_symptoms, severity_severe_non_severe) %>%
  summarise(numerator = sum(symptom_numerator, na.rm = TRUE), 
            denominator = sum(denominator, na.rm = TRUE)) %>%
  mutate(proportion_symptom= numerator/sum(denominator),
         proportion_symptom_low = Hmisc::binconf(numerator,sum(denominator))[,2],
         proportion_symptom_high = Hmisc::binconf(numerator,sum(denominator))[,3]) %>%
  ungroup() 

t <- tmp %>%  
  filter(severity_severe_non_severe == "non-severe") %>%
  group_by(grouped_symptoms) %>%
  summarise(proportion_symptom = sum(proportion_symptom, na.rm = T))%>%
  arrange(-proportion_symptom)


tmp %<>%
  mutate(grouped_symptoms=factor(tmp$grouped_symptoms, 
                                 levels = c(t$grouped_symptoms,
                                            unique(tmp$grouped_symptoms)[
                                              which(!unique(tmp$grouped_symptoms) %in%
                                                      t$grouped_symptoms)
                                            ])))
                                

tmp %>% 
  ggplot() + 
  geom_col(aes(x = grouped_symptoms,
               y = proportion_symptom,
               fill = grouped_symptoms)) +
  geom_errorbar(aes(x = grouped_symptoms,
                    ymin = proportion_symptom_low,
                    ymax = proportion_symptom_high))+
  theme_pubr() +
  scale_fill_viridis_d()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        legend.position = "none") +
  labs(x = "Symptoms", y = "Proportion of all reported symptoms") +
  facet_wrap(~severity_severe_non_severe, ncol = 1, 
             labeller = labeller(severity_severe_non_severe = setNames(
               c("Unspecified", "Non-severe", "Severe"),
               unique(tmp$severity_severe_non_severe))
             ),
             scales = "free_y")


```





```{r proportional_comorbidity_all, fig.cap="Proportion of all comorbidities of each type."}
# comorbidity
df_symptom %>%
  filter(!is.na(grouped_comorbidity)) %>%
  group_by(grouped_comorbidity) %>%
  summarise(numerator = sum(comorbidity_numerator, na.rm = TRUE), 
            denominator = sum(cohort_size, na.rm = TRUE)) %>%
  mutate(proportion_comorbidity= numerator/denominator,
         proportion_comorbidity_low = Hmisc::binconf(numerator,denominator)[,2],
         proportion_comorbidity_high = Hmisc::binconf(numerator,denominator)[,3])%>%
  
  ggplot() + 
  geom_col(aes(x = reorder(grouped_comorbidity, -proportion_comorbidity), 
               y = proportion_comorbidity,
               fill = proportion_comorbidity)) +
      geom_errorbar(aes(x = reorder(grouped_comorbidity, -proportion_comorbidity),
                    ymin = proportion_comorbidity_low,
                    ymax = proportion_comorbidity_high))+
  theme_pubr() +
  scale_fill_viridis_c()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        legend.position = "none") +
  labs(x = "Comorbidity", y = "Proportion of all reported comorbidity")
```

```{r proportional_comorbidity_by_severity, fig.cap="Proportion of all comorbidities structured by severity."}
# comorbidity

tmp <- df_symptom %>%
  filter(!is.na(grouped_comorbidity)) %>%
  group_by(grouped_comorbidity, severity_severe_non_severe) %>%
  summarise(numerator = sum(comorbidity_numerator, na.rm = TRUE), 
            denominator = sum(cohort_size, na.rm = TRUE)) %>%
  mutate(proportion_comorbidity= numerator/sum(denominator),
         proportion_comorbidity_low = Hmisc::binconf(numerator,sum(denominator))[,2],
         proportion_comorbidity_high = Hmisc::binconf(numerator,sum(denominator))[,3]) %>%
  ungroup() 


t <- tmp %>%  
  filter(severity_severe_non_severe == "non-severe") %>%
  group_by(grouped_comorbidity) %>%
  summarise(proportion_comorbidity = sum(proportion_comorbidity))%>%
  arrange(-proportion_comorbidity)


tmp %<>%
  mutate(grouped_comorbidity=factor(tmp$grouped_comorbidity, 
                                 levels = c(t$grouped_comorbidity,
                                            unique(tmp$grouped_comorbidity)[
                                              which(!unique(tmp$grouped_comorbidity) %in%
                                                      t$grouped_comorbidity)
                                            ])))
                                

tmp %>% 
  ggplot() + 
  geom_col(aes(x = grouped_comorbidity,
               y = proportion_comorbidity,
               fill = grouped_comorbidity)) +
  geom_errorbar(aes(x = grouped_comorbidity,
                    ymin = proportion_comorbidity_low,
                    ymax = proportion_comorbidity_high))+
  theme_pubr() +
  scale_fill_viridis_d()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        legend.position = "none") +
  labs(x = "Comorbidity", y = "Proportion of all reported comorbidities") +
  facet_wrap(~severity_severe_non_severe, ncol = 1, 
             labeller = labeller(severity_severe_non_severe = setNames(
               c("Unspecified", "Non-severe", "Severe"),
               unique(tmp$severity_severe_non_severe))
             ),
             scales = "free_y")

```

These figures depend on the number of non-empty records. We find that there are `r df_symptom %>% filter(!is.na(grouped_symptoms)) %>% select(symptom_numerator) %>% sum(na.rm = TRUE)` people reporting symptoms out of `r df_symptom %>% filter(!is.na(grouped_symptoms)) %>% select(denominator) %>% sum(na.rm = TRUE)` and there are `r df_symptom %>% filter(!is.na(grouped_comorbidity)) %>% select(comorbidity_numerator) %>% sum(na.rm = TRUE)` people reporting comorbidities out of `r df_symptom %>% filter(!is.na(grouped_comorbidity)) %>% select(denominator) %>% sum(na.rm = TRUE)`. 

In the googlesheet, `r df_symptom %>% filter(severity_severe_non_severe == "severe" ) %>% select(denominator) %>% sum(na.rm = TRUE)` people report severe disease, `r df_symptom %>% filter(severity_severe_non_severe == "non-severe" ) %>% select(denominator) %>% sum(na.rm = TRUE)` people report non-severe disease and `r df_symptom %>% filter(severity_severe_non_severe == "any_severity" ) %>% select(denominator) %>% sum(na.rm = TRUE)` do not specify.

# Other characteristics

## Age

```{r median_age_distribution, fig.cap="Age distribution of clinical studies"}
df_symptom %<>%
  mutate(median_age = as.numeric(median_age))


```

The average age was `r weighted.mean(df_symptom$median_age,w = df_symptom$cohort_size, na.rm = T)`.


```{r over_60}
df_symptom %<>%
  mutate(over_65_numerator = ifelse(is.na(over_65_numerator),
                                    over_60_numerator, 
                                    over_65_numerator))
df_over60 <- df_symptom %>% filter(!is.na(over_65_numerator))

```

The number of people who are at least 60, where reported, is `r df_over60 %>% select(over_65_numerator) %>% sum(na.rm = TRUE)` out of `r df_over60 %>% select(denominator) %>% sum(na.rm = TRUE)`. This is an underestimate as some studies only report people over 65.

## Gender

The number of people who are male, where reported, is `r df_symptom %>% filter(!is.na(male_numerator)) %>% select(male_numerator) %>% sum(na.rm = TRUE)` out of `r df_symptom %>% filter(!is.na(male_numerator)) %>% select(denominator) %>% sum(na.rm = TRUE)`.

# Pneumonia

Now we look specifically at the occurrence of pneumonia and who gets it.

```{r pneumonia}

df_pneumonia <- df_symptom %>% filter(grouped_symptoms %in% "pneumonia")

```

Out of people reporting pneumonia,`r df_pneumonia %>% filter(!is.na(male_numerator)) %>% select(male_numerator) %>% sum(na.rm = TRUE)` are male out of `r df_pneumonia %>% filter(!is.na(male_numerator)) %>% select(denominator) %>% sum(na.rm = TRUE)` reports and `r df_pneumonia %>% filter(!is.na(over_65_numerator)) %>% select(over_65_numerator) %>% sum(na.rm = TRUE)` are over the age of 60 from the pool of `r df_pneumonia %>% filter(!is.na(over_65_numerator)) %>% select(denominator) %>% sum(na.rm = TRUE)` completed reports.

The average age was `r weighted.mean(df_pneumonia$median_age,w = df_pneumonia$cohort_size, na.rm = T)`.

# Hubei

Now we look specifically at those cases reported to be in Hubei. In most cases this is specifically Wuhan.

```{r hubei}

df_hubei <- df_symptom %>% filter(!is.na(hubei_numerator))

```

Out of people reporting being from Hubei, `r df_hubei %>% filter(!is.na(male_numerator)) %>% select(male_numerator) %>% sum(na.rm = TRUE)` are male out of `r df_hubei %>% filter(!is.na(male_numerator)) %>% select(denominator) %>% sum(na.rm = TRUE)` reports and `r df_hubei %>% filter(!is.na(over_65_numerator)) %>% select(male_numerator) %>% sum(na.rm = TRUE)` are over the age of 60 from a pool of `r df_hubei %>% filter(!is.na(over_65_numerator)) %>% select(denominator) %>% sum(na.rm = TRUE)` reports.

The average age was `r weighted.mean(df_hubei$median_age,w = df_hubei$cohort_size, na.rm = T)`.


