---
title: "import severity data"
author: "Natsuko Imai and Katy Gaythorpe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_width: 7
    fig_height: 5
    fig_caption: true
    highlight: "tango"
    reference_docx: word_style.docx
    df_print: "kable"
---

# Cleaning China papers and preprints

```{r set up }

 knitr::opts_chunk$set(echo = FALSE)

## library(orderly)
## orderly::orderly_test_start("B1_clean_severity_data")
## orderly::orderly_run("B1_clean_severity_data")

```

Import data and change numeric columns.
```{r China_papers 1 }

## Read directly from google drive to avoid having to download the
## data. Note that this step will prompt for google sign in and
## ask you to allow Tidyverse API to access your srive.
googlesheets4::sheets_deauth()
df_raw <- googlesheets4::read_sheet(
  "https://docs.google.com/spreadsheets/d/1jBbIdYObHrIWz4xaQRDXVFEQE-NbcqiL7DOjdSR39-o/edit?usp=sharing",
  sheet = "china_papers_preprints",
  col_types = "c", ## read everything as character, easier than having to guess excel's date format
  na = c("", "NA", " ") ## sometimes people actually write NA rather than leaving a cell blank.
)

df <- janitor::clean_names(df_raw) %>% as.data.frame()

paste0("Number of entries: ", nrow(df))

numeric_col_indices <- grep("numerator|denominator|size", names(df))
df[,numeric_col_indices] <-  sapply(numeric_col_indices, 
       FUN = function(x){
         df[,x] <- as.numeric(df[,x])
       })

paste0("Numeric columns found: ", paste0(names(df)[numeric_col_indices], collapse = ", "))

# only want clinical papers
df %<>% dplyr::filter(manuscript_type=="CLINICAL")

```


Cleaning symptoms and comorbidities, this uses `symptom_meta` and `comorbidity_meta` which may need updating if symptoms are unmatched.

```{r China papers 2}

# cleaning symptoms and comorbidities

# start grouping things and get rid of capitals
df %<>% mutate(symptom = tolower(symptom),
                       comorbidity = tolower(comorbidity))

# now satrt grouping symptoms
symptom_meta <- read.csv("symptom_meta.csv", stringsAsFactors = FALSE)

if(sum(!df$symptom %in% symptom_meta$symptom)>0){
  stop("Update symptom meta, there are symptoms that are not currently mapped.")
}

df$grouped_symptoms <- symptom_meta$symptom_group[match(df$symptom, symptom_meta$symptom)]

comorbidity_meta <- read.csv("comorbidity_meta.csv", stringsAsFactors = FALSE)

if(sum(!df$comorbidity %in% comorbidity_meta$comorbidity)>0){
  stop("Update comorbidity meta, there are comorbidities that are not currently mapped.")
}

df$grouped_comorbidity <- comorbidity_meta$comorbidity_group[match(df$comorbidity,  comorbidity_meta$comorbidity)]

```

Save the resulting raw and processed data.

```{r China papers 3}
# save the output
readr::write_csv(df_raw, "severity_data_China_papers_raw.csv")
readr::write_csv(df, "severity_data_China_papers.csv")
```


# Cleaning country case studies

Importing data.

```{r case-studies-1 }
## Read directly from google drive to avoid having to download the
## data. Note that this step will prompt for google sign in and
## ask you to allow Tidyverse API to access your srive.
googlesheets4::sheets_deauth()
df_raw <- googlesheets4::read_sheet(
  "https://docs.google.com/spreadsheets/d/1jBbIdYObHrIWz4xaQRDXVFEQE-NbcqiL7DOjdSR39-o/edit?usp=sharing",
  sheet = "country_case_studies",
  col_types = "c", ## read everything as character, easier than having to guess excel's date format
  na = c("", "NA", " ") ## sometimes people actually write NA rather than leaving a cell blank.
)

df <- janitor::clean_names(df_raw) %>% as.data.frame()

paste0("Number of entries: ", nrow(df))


```

```{r check_ids}
if(sum(duplicated(df$new_id))>0){
  stop(paste0("There are duplicated ids, look at: ", paste0(df$new_id[duplicated(df$new_id)], 
                                                                      collapse = ", ") ))
}

```

Fixing the date columns, in this week look for column names beginning with `date`.

```{r case-studies-2}
# fix dates etc.
date_col_indices <- grep("date", names(df))
for(x in date_col_indices){
  df[,x] <- lubridate::dmy(df[,x], quiet = TRUE)
}

paste0("Date columns found: ", paste0(names(df)[date_col_indices], collapse = ", "))
```

Fixing the age values. Most ages are either stated exactly of as decades ie. "40s", in the latter we change them to the middle of the decade ie. "45".

```{r case-studies-3}
# fix age
# decades
df %<>% mutate(age = ifelse(grepl("s", age),
                            as.numeric(gsub("s", "", age)) + 5,
                            age))

# less than
df %<>% mutate(age = ifelse(grepl("<", age),
                            as.numeric(gsub("<", "", age))/2,
                            age))

# elderly
df$age[df$age == "elderly"] = 75


df %<>% mutate(age = as.numeric(age))
   
unique(df$age) 


df %<>% mutate(over_60 = ifelse(age>59, 1,0) )

ggplot(df) + 
  geom_histogram(aes(x = age, y = ..density..), binwidth = 10) +
  theme_pubr()
    
```

Now, we fix the symptoms. In this we produce 3 tables, symptoms for everyone at each stage and then divided by the over and under 60s.

```{r case-studies-4}
# check the symptoms
symptom_col_indices <- grep("^symptoms_at_onset|^symp_prog1|^symp_prog2|^symp_prog3|^symp_prog4|^any_symp",
                            names(df))

df_symp <-  df %>% mutate_at(names(df)[symptom_col_indices],
                  strsplit, split = ";")

paste0("Symptom columns found: ", 
       paste0(names(df_symp)[symptom_col_indices], collapse = ", "))




```

Translate symptoms in main file.

```{r case-studies-8 }

for( i in 1:length(symptom_col_indices)){
  new_col = ncol(df_symp)+1
  df_symp[, new_col] = rep(NA, nrow(df_symp))
  
  for(j in 1:nrow(df_symp)){
    if(all(unlist(df_symp[j, symptom_col_indices[i]]) %in% symptom_meta$symptom)){
      
      
      tmp1 <- (symptom_meta$symptom_group[match(unlist(df_symp[j, 
                                                               symptom_col_indices[i] ]), 
                                                symptom_meta$symptom) ])
      
      df_symp[j, symptom_col_indices[i] ][[1]] <- list(tmp1)
      
      #add severity
      
      tmp2 <- (symptom_meta$severity_group[match(unlist(df_symp[j, 
                                                                symptom_col_indices[i] ]), 
                                                 symptom_meta$symptom) ])
      
      df_symp[j, new_col][[1]] <- list(tmp2)
      
    } else {
      stop(paste0("A symptom is missing- something in: ", 
                  paste0(unlist(df_symp[j, symptom_col_indices[i] ]), collapse = ", ")))
    }
    
  }
}

# sort out severity
df_symp$severity = rep("mild", nrow(df_symp))
severity_col_indices = grep("^V", names(df_symp))
for(j in 1:nrow(df_symp)){
  for(i in severity_col_indices){
  if("severe" %in% unlist(df_symp[j,i]) ){
    df_symp$severity[j] = "severe"
  } 
  }
}
df_symp$severity[is.na(df_symp$severity)] = "mild"

df_symp %<>% dplyr::select(-starts_with("V"))

# if a person died then overwrite severity
df_symp %<>% mutate(severity = ifelse(!is.na(date_death),
                                      "severe",
                                      severity))

table(df_symp$severity)


```

All symptoms.

```{r case-studies-5 }

# recode symptoms
symptoms_df_all <- get_symptoms(df_symp, symptom_meta)
symptoms_df_over_60 <- get_symptoms(df_symp %>% filter(over_60==1), 
                                    symptom_meta)
symptoms_df_under_60 <- get_symptoms(df_symp %>% filter(over_60==0), 
                                    symptom_meta)

symptoms_df_all
```

Symptoms in over 60s.

```{r case-studies-6 }
symptoms_df_over_60
```

Symptoms in under 60s.

```{r case-studies-7 }
symptoms_df_under_60
```

# Severity

We consider the severity of each case based on their symptoms and outcome ie. those who died are considered severe infections.




Save all files.

```{r case-studies-9 }

readr::write_csv(x = df_raw, path = "severity_data_country_case_studies_raw.csv")

saveRDS(df_symp, "severity_data_country_case_studies.rds")

readr::write_csv(x = symptoms_df_all, 
                 path = "severity_data_case_study_symptoms_all.csv")
readr::write_csv(x = symptoms_df_over_60, 
                 path = "severity_data_case_study_symptoms_over_60.csv")
readr::write_csv(x = symptoms_df_under_60, 
                 path = "severity_data_case_study_symptoms_under_60.csv")


```
