---
title: "Plot severity from case studies"
author: "Katy Gaythorpe and Natsuko Imai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_width: 7
    fig_height: 5
    fig_caption: true
    highlight: "tango"
    reference_docx: word_style.docx
    df_print: "kable"
---

```{r set up }

 knitr::opts_chunk$set(echo = FALSE, fig.path = "figures/", 
                       fig.ext = "png",
                       fig.height = 8)

## library(orderly)
## orderly::orderly_test_start("B3_plot_severity_case_studies")
## orderly::orderly_run("B3_plot_severity_case_studies")


## orderly::orderly_develop_start()
## orderly::orderly_develop_status()
## orderly::orderly_develop_clean()

```

Import data.
```{r }

symptoms_freq <- read.csv("severity_data_case_study_symptoms_all.csv", 
                          stringsAsFactors = FALSE,
                          strip.white = TRUE)
symptoms_freq_over60 <- read.csv("severity_data_case_study_symptoms_over_60.csv", 
                          stringsAsFactors = FALSE,
                          strip.white = TRUE)
symptoms_freq_under60 <- read.csv("severity_data_case_study_symptoms_under_60.csv", 
                          stringsAsFactors = FALSE,
                          strip.white = TRUE)

df <- readRDS("severity_data_country_case_studies.rds")

```

# Simple symptom progression

```{r symptoms_all}

symptoms_freq_tmp <-  symptoms_freq %>% 
  filter(!symptom %in% c("any_symptom", "stable") ) # get specific symptoms only

# get proportion. Ignore any symptom not coded and NAs.
for(i in 2:ncol(symptoms_freq_tmp)){
  total_symp <- sum(symptoms_freq_tmp[,i], na.rm = TRUE)
  symptoms_freq_tmp[,i] <- symptoms_freq_tmp[,i]/total_symp
  symptoms_freq_tmp[nrow(symptoms_freq_tmp)+1, i] <- total_symp
}

# wide to long
symptoms_freq_prop <- symptoms_freq_tmp %>% 
  tidyr::gather(Progression, Frequency, -symptom)

# calc freq of each symptom amongst those who report it
# not total number of patients
# does not have to some to one.
symptoms_freq_prop %>% 
  filter(!symptom %in% c("Total reported", NA) ) %>%
  filter(!is.na(Frequency)) %>%
  mutate(Frequency = as.numeric(Frequency)) %>%
  group_by(symptom) %>%
  ungroup() %>%
  
  ggplot() + 
  geom_col(aes(x=symptom, 
               y = Frequency, 
               fill = Progression))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 14))+
  ylab("Proportion") + xlab("Symptom") +
  scale_fill_manual(labels = c(freq_at_any = "Any",
                               freq_at_onset = "Onset",
                               freq_at_prog1 = "Progression 1",
                               freq_at_prog2 = "Progression 2",
                               freq_at_prog3 = "Progression 3",
                               freq_at_prog4 = "Progression 4"),
                    values=c(rev(wes_palette("Darjeeling2")), "orchid") )

```

```{r symptoms_over60}

# same as above but subset by >60 years
symptoms_freq_over60_tmp <- symptoms_freq_over60 %>% 
  filter(!symptom %in% c("any_symptom", "stable") ) # get specific symptoms only

for(i in 2:ncol(symptoms_freq_over60_tmp)){
  total_symp<- sum(symptoms_freq_over60_tmp[,i], na.rm = TRUE)
  symptoms_freq_over60_tmp[,i] <- symptoms_freq_over60_tmp[,i]/total_symp
  symptoms_freq_over60_tmp[nrow(symptoms_freq_over60_tmp)+1, i] <- total_symp
}

symptoms_freq_over60_prop <- symptoms_freq_over60_tmp %>% 
  tidyr::gather(Progression, Frequency, -symptom)


symptoms_freq_over60_prop %>% 
  filter(!symptom %in% c("Total reported", NA) ) %>%
  filter(!is.na(Frequency)) %>%
  mutate(Frequency = as.numeric(Frequency)) %>%
  group_by(symptom) %>%
  ungroup() %>%
  
  ggplot() + 
  geom_col(aes(x=symptom, 
               y = Frequency, 
               fill = Progression))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1,size = 14))+
  ylab("Proportion") + xlab("Symptom") +
  scale_fill_manual(labels = c(freq_at_any = "Any",
                                  freq_at_onset = "Onset",
                                  freq_at_prog1 = "Progression 1",
                                  freq_at_prog2 = "Progression 2",
                                  freq_at_prog3 = "Progression 3",
                                  freq_at_prog4 = "Progression 4"),
                       values=c(rev(wes_palette("Darjeeling2")), "orchid") )

```

```{r symptoms_under60}
# same as above but subset by <60 years

symptoms_freq_under60_tmp <- symptoms_freq_under60 %>% 
  filter(!symptom %in% c("any_symptom", "stable") ) # get specific symptoms only

for(i in 2:ncol(symptoms_freq_under60_tmp)){
  total_symp<- sum(symptoms_freq_under60_tmp[,i], na.rm = TRUE)
  symptoms_freq_under60_tmp[,i] <- symptoms_freq_under60_tmp[,i]/total_symp
  symptoms_freq_under60_tmp[nrow(symptoms_freq_under60_tmp)+1, i] <- total_symp
}

symptoms_freq_under60_prop <- symptoms_freq_under60_tmp %>% 
  tidyr::gather(Progression, Frequency, -symptom)


symptoms_freq_under60_prop %>% 
  filter(!symptom %in% c("Total reported", NA) ) %>%
  filter(!is.na(Frequency)) %>%
  mutate(Frequency = as.numeric(Frequency)) %>%
  group_by(symptom) %>%
  ungroup() %>%
  
  ggplot() + 
  geom_col(aes(x=symptom, 
               y = Frequency, 
               fill = Progression))+
  theme_pubr()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 14))+
  ylab("Proportion") + xlab("Symptom") +
  scale_fill_manual(labels = c(freq_at_any = "Any",
                                  freq_at_onset = "Onset",
                                  freq_at_prog1 = "Progression 1",
                                  freq_at_prog2 = "Progression 2",
                                  freq_at_prog3 = "Progression 3",
                                  freq_at_prog4 = "Progression 4"),
                       values=c(rev(wes_palette("Darjeeling2")), "orchid") )

```

# Progression map of all individuals

```{r progressionmap_all, fig.height=12}
make_progressionmap(df)


# dates wrong - could move to cleaning
if(min(df$date_hospitalised-df$date_onset<0, na.rm = T)<0){
  stop(paste0("Please check hospitalisation date for: ",
              paste0(df$new_id[which(df$date_hospitalised-df$date_onset<0)], 
                                                                collapse=", ")))
}

```

```{r progressionmap_over60, fig.height=12}

make_progressionmap(df %>% filter(as.numeric(age)>=60))


```

```{r progressionmap_under60, fig.height=12}

make_progressionmap(df %>% filter(as.numeric(age)<=60))


```


# Time to symptoms

## Pneumonia

```{r time_pneumonia, fig.height=6}
df <- add_first_symptom(df, "pneumonia")

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "cyan") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to pneumonia") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}
fit

if(fit$converged == FALSE){
  paste0("check fit to dist. as has not converged")
}

mean_onset_to_symp <- data.frame(symptom = "pneumonia",
                                 mean_onset_to_symp = fit$mu,
                                 sd_onset_to_symp = fit$sd)
```

We see pneumonia reported

```{r pneumonia at onset}

# what proportion of all pneumonia at is at onset
df_pneumonia <- df %>% 
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_pneumonia$name_first))

```
out of `r nrow(df_pneumonia)` reports where pneumonia is mentioned. 

Of those who have recovered ...

```{r pneumonia at onset recovered}

# what proportion of all pneumonia at is at onset
df_pneumonia <- df %>% 
  filter(!is.na(date_hospitalised)) %>%
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_pneumonia$name_first))

```
out of `r nrow(df_pneumonia)` reports where pneumonia is mentioned and the person recovered.

## Fever

```{r time_fever, fig.height=6}
df <- add_first_symptom(df, "fever")

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "orange") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to fever") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}
fit

if(fit$converged == FALSE){
  paste0("check fit to dist. as has not converged")
}

mean_onset_to_symp %<>% bind_rows( data.frame(symptom = "fever",
                                 mean_onset_to_symp = fit$mu,
                                 sd_onset_to_symp = fit$sd) )

```

We see fever reported

```{r fever at onset}

# what proportion of all pneumonia at is at onset
df_p <- df %>% 
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_p$name_first))

```
out of `r nrow(df_p)` reports. 

## Shortness of breath

```{r time_breath, fig.height=6}
df <- add_first_symptom(df, "respiratory symptoms")

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "pink") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to respiratory symptoms") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}
fit
if(fit$converged == FALSE){
  paste0("check fit to dist. as has not converged")
}

mean_onset_to_symp %<>% bind_rows( data.frame(symptom = "respiratory symptoms",
                                 mean_onset_to_symp = fit$mu,
                                 sd_onset_to_symp = fit$sd) )
```

We see respiratory symptoms reported

```{r breath at onset}

# what proportion of all respiratory symptoms is at onset
df_p <- df %>% 
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_p$name_first))

```
out of `r nrow(df_p)` reports. 

## Diarrhea

```{r time_diarrhea, fig.height=6}
df <- add_first_symptom(df, "diarrhea")

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "brown") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to diarrhea") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}
fit
if(fit$converged == FALSE){
  paste0("check fit to dist. as has not converged")
}

mean_onset_to_symp %<>% bind_rows( data.frame(symptom = "diarrhea",
                                 mean_onset_to_symp = fit$mu,
                                 sd_onset_to_symp = fit$sd) )
```

We see diarrhea reported

```{r diarrhea at onset}

# what proportion of all diarrhoea at is at onset
df_p <- df %>% 
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_p$name_first))

```
out of `r nrow(df_p)` reports. 

## Fatigue

```{r time_fatigue, fig.height=6}
df <- add_first_symptom(df, "myalgia or fatigue")

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "green") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to myalgia or fatigue") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}

fit
if(fit$converged == FALSE){
  paste0("check fit to dist. as has not converged")
}

mean_onset_to_symp %<>% bind_rows( data.frame(symptom = "myalgia or fatigue",
                                 mean_onset_to_symp = fit$mu,
                                 sd_onset_to_symp = fit$sd) )

```

We see myalgia or fatigue reported

```{r fatigue at onset}

# what proportion of all fatigue at is at onset
df_p <- df %>% 
  filter(!is.na(name_first)) %>% 
  filter(name_first != "date_symptoms_full")

as.data.frame(table(df_p$name_first))

```
out of `r nrow(df_p)` reports. 

# Number of reported symptoms

```{r any symptoms}

symptom_col_indices <- 
  grep("^symptoms_at_onset|^symp_prog1|^symp_prog2|^symp_prog3|^symp_prog4|any_symp",
                            names(df))
# number reporting any symptoms
df %<>% mutate(symptom_reported = NA)

for(i in 1:nrow(df)){
  tmp=df[i, symptom_col_indices]
  tmp[tmp=="any_symptom"]=NA
  df$symptom_reported[i] = all(is.na(tmp))
}

```

The number of people reporting symptoms in the data is `r sum(!df$symptom_reported)`.

# Time to outcome

## Recovery or discharge

```{r time_recovery, fig.height=6}
df %<>% mutate(time_first = date_recovered-date_onset)

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "darkgreen") +
  geom_line(data = data.frame(x = seq(0,50, by = 0.1),
                              y = dgamma(seq(0,50, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 50) +
  theme_pubr()+
  xlab("Time from onset of symptoms to recovery or discharge") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>50)], 
                                        collapse = ", "))
}
fit

if(fit$converged == FALSE){
  paste0("check your gamma fit as it has not converged")
}
```


## Hospitalised

```{r time_hospitalised, fig.height=6}
df %<>% mutate(time_first = date_hospitalised-date_onset)

#ignore previous hospitalisation for now- japan052
df %<>% mutate(time_first = ifelse(time_first<0, NA, time_first))

fit <- epitrix::fit_disc_gamma(as.numeric(df$time_first))

ggplot(df) + 
  geom_histogram(aes(x = time_first, 
                     y = ..count../sum(..count..)), fill = "blue") +
  geom_line(data = data.frame(x = seq(0,20, by = 0.1),
                              y = dgamma(seq(0,20, by = 0.1),
                                         shape = fit$distribution$parameters$shape,
                           scale = fit$distribution$parameters$scale)),
            aes(x = x, y = y))+
  
  xlim(-1, 20) +
  theme_pubr()+
  xlab("Time from onset of symptoms to hospitalised") 

dates_wrong <- sum(df$time_first<0 | df$time_first>20, 
                   na.rm = TRUE)

if(dates_wrong>0){
  paste0("Check the dates for ", paste0(df$new_id[which(df$time_first<0 |
                                          df$time_first>20)], 
                                        collapse = ", "))
}
fit
if(fit$converged == FALSE){
  paste0("check your gamma fit as it has not converged")
}

```

# Age

```{r age}
# TO DO: plot by 10 yr age bands
df %>%
  ggplot()+
  geom_histogram(aes(age, fill = pneumonia))+
  theme_pubr() +
  xlab("Age")+ ylab("Count") +
  geom_vline(xintercept = 60)
```

```{r symptom_by_age}

symp_by_age <- bind_rows( symptoms_freq_under60 %>% mutate(over60 = "under60"),
                          symptoms_freq_over60 %>% mutate(over60 = "over60") )

symp_by_age$total_symp <- rowSums(symp_by_age[,grep("^freq_", 
                                                    names(symp_by_age))], na.rm = TRUE)

symp_by_age %>%
  filter(!symptom %in% c("Total reported", NA, "any_symptom") ) %>%
  group_by(symptom) %>%
  mutate(count_by_symptom = sum(total_symp))%>%
  ggplot() +
  geom_col(aes(x = symptom, 
               y = total_symp/count_by_symptom, 
               fill = as.factor(over60))) +
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c(wes_palette("Darjeeling2")[5],
                               wes_palette("Darjeeling2")[1]),
                    labels = c("over60" = "Over 60",
                               "under60" = "Under 60"),
                    name = "") +
  ylab("Proportion") + xlab("")

```


# Tables of outputs

```{r table outputs 1}

# symptoms table
table_out <- data.frame(symptom = unique(symp_by_age$symptom))

table_out %<>% filter(!symptom %in% c("any_symptom", "no symptoms", "stable"))

# all
table_out %<>% left_join(symp_by_age %>%
                           group_by(symptom) %>%
                           mutate(count_by_symptom = sum(total_symp)) %>%
                           select(c(symptom, count_by_symptom)))
table_out %<>% group_by(symptom) %>% summarise(proportion_of_all = count_by_symptom[1])

table_out %<>% 
  mutate(proportion_of_all = proportion_of_all/sum(proportion_of_all, na.rm=TRUE))

#over 60
table_out %<>% left_join(symp_by_age %>%
                           filter(over60 == "over60") %>%
                           select(c(symptom, total_symp)))
table_out %<>% 
  mutate(total_symp = total_symp/sum(total_symp, na.rm=TRUE))

table_out %<>% rename(proportion_of_over_60 = total_symp)

#under 60
table_out %<>% left_join(symp_by_age %>%
                           filter(over60 == "under60") %>%
                           select(c(symptom, total_symp)))
table_out %<>% 
  mutate(total_symp = total_symp/sum(total_symp, na.rm=TRUE))

table_out %<>% rename(proportion_of_under_60 = total_symp)

# Total

table_out %<>% bind_rows(data.frame(symptom = "Total reported",
                                    proportion_of_all = sum(
                                      symp_by_age[ !symp_by_age$symptom %in% c("any_symptom", "no symptoms", "stable"), 
                                                  grep("^freq", names(symp_by_age))][],
                                      na.rm = TRUE
                                    ),
                                    proportion_of_over_60 = sum(
                                      symp_by_age[ !symp_by_age$symptom %in% c("any_symptom", "no symptoms", "stable") &
                                                     symp_by_age$over60 == "over60",
                                                   
                                                   grep("^freq", names(symp_by_age))][],
                                      na.rm = TRUE
                                    ),
                                    proportion_of_under_60 = sum(
                                      symp_by_age[!symp_by_age$symptom %in% c("any_symptom", "no symptoms", "stable") &
                                                  symp_by_age$over60 == "under60",
                                                   grep("^freq", names(symp_by_age))][],
                                      na.rm = TRUE
                                    )
))

# print
table_out %>% arrange(-proportion_of_all) %>% mutate_if(is.numeric, round, digits = 2)


#save
saveRDS(table_out, "proportional_symptoms.rds")

```


```{r table outputs 2}
table_onset_to_symp <- mean_onset_to_symp %>% 
  rename(fitted_mean = mean_onset_to_symp,
                               fitted_sd = sd_onset_to_symp)

table_onset_to_symp %<>% bind_rows(data.frame(symptom = unique(symptoms_freq$symptom),
                                             fitted_mean = NA,
                                             fitted_sd = NA))

table_onset_to_symp %<>% filter(!duplicated(symptom))

table_onset_to_symp %<>% filter(!symptom %in% c("any_symptom", "no symptoms", "stable"))

table_onset_to_symp %<>% mutate(observed_mean = NA,
                                observed_sd = NA)

for(i in 1:nrow(table_onset_to_symp)){
  out <- get_mean_onset_to_symptom(df, table_onset_to_symp$symptom[i])
  
  table_onset_to_symp$observed_mean[i] <- out$mean_out
  table_onset_to_symp$observed_sd[i] <- out$sd_out
}

# total
table_onset_to_symp %<>% left_join(symp_by_age %>%
                           group_by(symptom) %>%
                           mutate(count_by_symptom = sum(total_symp)) %>%
                           select(c(symptom, count_by_symptom))) %>%
  unique() %>%
  rename(number_of_symptom_reports = count_by_symptom)

#print
table_onset_to_symp %>% arrange(-observed_mean) %>% mutate_if(is.numeric, round, digits = 2)

#save
saveRDS(table_onset_to_symp, "time_onset_to_symptom.rds")
```


```{r table outputs 3}

comb_table <- left_join(table_out, table_onset_to_symp)

comb_table %<>% select(-starts_with("fitted"))

comb_table %<>% rename(`Mean time from onset to symptom` = observed_mean,
                       `S.D. time from onset to symptom` = observed_sd,
                       Symptom = symptom,
                       `Proportion of all reported symptoms` = proportion_of_all,
                       `Proportion of reported symptoms in over 60s` = proportion_of_over_60,
                       `Proportion of reported symptoms in under 60s` = proportion_of_under_60,
                       `Total count of individual symptom` = number_of_symptom_reports)

comb_table

saveRDS(comb_table, "combined_case_study_table.rds")
```

```{r table outputs 4}
## these are all observed mean and sd delays.
## TO DO: repeat for fitted delays and check how diff they are.

table_onset_to_outcome <- data.frame(Outcome = c("First consultation", "Hospitalised", "Recovered/Discharged", "Death", "Hospitalised duration"),
                                     Mean = rep(NA, 5),
                                     Sd = rep(NA, 5),
                                     Number_reports = rep(NA, 5))

# first consultation
table_onset_to_outcome$Mean[table_onset_to_outcome$Outcome == "First consultation"] <- 
  mean(df$date_first_consultation - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Sd[table_onset_to_outcome$Outcome == "First consultation"] <- 
  sd(df$date_first_consultation - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Number_reports[table_onset_to_outcome$Outcome == "First consultation"] <- 
  sum(!is.na(df$date_first_consultation))

# hospitalised
df %<>% mutate(time_first = date_hospitalised - date_onset)
df %<>% mutate(time_first = ifelse(time_first<0, NA, time_first)) #remove negative ones for now

table_onset_to_outcome$Mean[table_onset_to_outcome$Outcome == "Hospitalised"] <- 
  mean(df$date_hospitalised - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Sd[table_onset_to_outcome$Outcome == "Hospitalised"] <- 
  sd(df$date_hospitalised - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Number_reports[table_onset_to_outcome$Outcome == "Hospitalised"] <- 
  sum(!is.na(df$date_hospitalised))

# recovered
table_onset_to_outcome$Mean[table_onset_to_outcome$Outcome == "Recovered/Discharged"] <- 
  mean(df$date_recovered - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Sd[table_onset_to_outcome$Outcome == "Recovered/Discharged"] <- 
  sd(df$date_recovered - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Number_reports[table_onset_to_outcome$Outcome == "Recovered/Discharged"] <- 
  sum(!is.na(df$date_recovered))

# death
table_onset_to_outcome$Mean[table_onset_to_outcome$Outcome == "Death"] <- 
  mean(df$date_death - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Sd[table_onset_to_outcome$Outcome == "Death"] <- 
  sd(df$date_death - df$date_onset, na.rm = TRUE)
table_onset_to_outcome$Number_reports[table_onset_to_outcome$Outcome == "Death"] <- 
  sum(!is.na(df$date_death))

# hospital stay
tmp=df$date_recovered - df$date_hospitalised
tmp2 = df$date_death - df$date_hospitalised
tmp[is.na(tmp)] = tmp2[is.na(tmp)]

table_onset_to_outcome$Mean[table_onset_to_outcome$Outcome == "Hospitalised duration"] <- 
  mean(tmp, na.rm = TRUE)
table_onset_to_outcome$Sd[table_onset_to_outcome$Outcome == "Hospitalised duration"] <- 
  sd(tmp, na.rm = TRUE)
table_onset_to_outcome$Number_reports[table_onset_to_outcome$Outcome == "Hospitalised duration"] <- 
  sum(!is.na(tmp))

table_onset_to_outcome %>% mutate_if(is.numeric, round, digits = 2)

#save
saveRDS(table_onset_to_outcome, "time_onset_to_outcome.rds")

```